### 2021-01-11 material/expstimuli/selection/analysis(3sd).Rmd

---
title: "Music Perception: Stimuli profiles (3sd)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

I chose stimuli with the IOIs within 3sd randomly.

Note: SubNr 8 was excluded due to an experimental error, SubNr 16 was excluded due to a deviated tempo (see analysis for [teaching-v1.0](https://github.com/atsukotominaga/teaching-v1.0))

```{r setup, include = FALSE}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
```

```{r file, include = FALSE}
# create a list of data file names
lf_a_t <- list.files("./3sd/art_teaching", pattern = "txt")
lf_a_p <- list.files("./3sd/art_performing", pattern = "txt")
lf_d_t <- list.files("./3sd/dyn_teaching", pattern = "txt")
lf_d_p <- list.files("./3sd/dyn_performing", pattern = "txt")

# merge all data files into one
data <- data.table()
for (i in 1:length(lf_a_t)){
 data_i <- fread(file.path("./3sd/art_teaching", lf_a_t[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_a_p)){
 data_i <- fread(file.path("./3sd/art_performing", lf_a_p[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_d_t)){
 data_i <- fread(file.path("./3sd/dyn_teaching", lf_d_t[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}
for (i in 1:length(lf_d_p)){
 data_i <- fread(file.path("./3sd/dyn_performing", lf_d_p[i]), header = T, sep = ",", dec = ".")
 data_i$Instance <- i
 data <- rbind(data, data_i)
}

# define subcomponents
# For intervals
ls_legato <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_staccato <- list(c(9:15), c(25:31), c(34:40), c(50:56))
ls_forte <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_piano <- list(c(9:15), c(25:31), c(34:40), c(50:56))

# For each note (velocity only)
ls_legato_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_staccato_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))
ls_forte_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_piano_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))

# Define Skill Change (LtoS, FtoP)
change_1 <- c(8, 24, 49)
# Define Skill Change (StoL, PtoF)
change_2 <- c(16, 41, 57)
```

# IOIs
```{r ioi, include = FALSE}
data_onset <- data[Key_OnOff == 1]
data_offset <- data[Key_OnOff == 0]

data_onset$IOI <- diff(c(0, data_onset$TimeStamp))
dt_ioi <- data_onset[NoteNr != 17]
dt_ioi$Interval <- rep(1:66, nrow(dt_ioi)/66)

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}

ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(Instance, Condition, Skill)]
ioi_art <- ioi[Skill == "articulation", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
ioi_dyn <- ioi[Skill == "dynamics", .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill)]
```

## Summary Stats
### Individual
```{r ioi_stats_ind, echo = FALSE}
ioi
```

### Averaged
```{r ioi_stats_all, echo = FALSE}
ioi_art
ioi_dyn
```

## Plots
### Articulation
```{r ioi_art_plot, echo = FALSE}
p_ioi_art <- ggplot(ioi[Skill == "articulation"], aes(x = Instance, y = Mean, fill = Condition)) +
 geom_bar(stat = "identity", position = position_dodge()) +
 geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
               width = .2, position = position_dodge(.9)) +
 geom_hline(yintercept = 188, linetype = "dashed") +
 scale_x_continuous(breaks = seq(1:24)) +
 labs(y = "IOI (ms)", subtitle = "IOI: Each instance (Articulation)")
p_ioi_art

p_ioi_art_all <- ggplot(ioi_art, aes(x = Skill, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                 width = .2, position = position_dodge(.9)) +
   geom_hline(yintercept = 188, linetype = "dashed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Averaged (Articulation)")
p_ioi_art_all
```
### Dynamics
```{r ioi_dyn_plot, echo = FALSE}
p_ioi_dyn <- ggplot(ioi[Skill == "dynamics"], aes(x = Instance, y = Mean, fill = Condition)) +
 geom_bar(stat = "identity", position = position_dodge()) +
 geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
               width = .2, position = position_dodge(.9)) +
 geom_hline(yintercept = 188, linetype = "dashed") +
 scale_x_continuous(breaks = seq(1:24)) +
 labs(y = "IOI (ms)", subtitle = "IOI: Each instance (Dynamics)")
p_ioi_dyn

p_ioi_dyn_all <- ggplot(ioi_dyn, aes(x = Skill, y = Mean, fill = Condition)) +
   geom_bar(stat = "identity", position = position_dodge()) +
   geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
                 width = .2, position = position_dodge(.9)) +
   geom_hline(yintercept = 188, linetype = "dashed") +
   labs(y = "IOI (ms)", subtitle = "IOI: Averaged (Dynamics)")
p_ioi_dyn_all
```

# KOT
```{r kot, include = FALSE}
data_onset$KOT <- NA
for (row in 1:nrow(data_onset)){
if (row < nrow(data_onset)){
  data_onset$KOT[row+1] <- data_offset$TimeStamp[row] - data_onset$TimeStamp[row+1] # offset(n) - onset(n+1)
}
}

dt_kot <- data_onset[NoteNr != 17]
dt_kot$Interval <- rep(1:66, nrow(dt_kot)/66)

# assign Subcomponents
dt_kot$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
for (note in 1:length(ls_legato[[phrase]])){
  dt_kot[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
}
}
# Staccato
for (phrase in 1:length(ls_staccato)){
for (note in 1:length(ls_staccato[[phrase]])){
  dt_kot[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
}
}

# Forte
for (phrase in 1:length(ls_forte)){
for (note in 1:length(ls_forte[[phrase]])){
  dt_kot[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
}
}
# Piano
for (phrase in 1:length(ls_piano)){
for (note in 1:length(ls_piano[[phrase]])){
  dt_kot[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
}
}

# assign Subcomponent Change
for (number in change_1){
dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}

kot <- dt_kot[Subcomponent == "Legato" | Subcomponent == "Staccato", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(Instance, Condition, Skill, Subcomponent)]
kot_all <- kot[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Subcomponent)]
```

## Plots
```{r kot_plot, echo = FALSE}
p_kot <- ggplot(kot, aes(x = Instance, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
facet_grid(Subcomponent ~ .) +
labs(y = "KOT (ms)", subtitle = "KOT: Each instance (Articulation)")
p_kot

p_kot_all <- ggplot(kot_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
labs(y = "KOT (ms)", subtitle = "KOT: Averaged (Articulation)")
p_kot_all
```

## Each instance
```{r kot_stats_ind, echo = FALSE}
kot
```

## Averaged
```{r kot_stats_all, echo = FALSE}
kot_all
```

# KV
```{r kv, include = FALSE}
dt_vel <- data_onset
dt_vel$Note <- rep(1:67, nrow(dt_vel)/67)

# assign Subcomponents
# for each note
dt_vel$Subcomponent <- NA
# Legato
for (phrase in 1:length(ls_legato_2)){
for (note in 1:length(ls_legato_2[[phrase]])){
  dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_legato_2[[phrase]][note]] <- "Legato"
}
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
for (note in 1:length(ls_staccato_2[[phrase]])){
  dt_vel$Subcomponent[dt_vel$Skill == "articulation" & dt_vel$Note == ls_staccato_2[[phrase]][note]] <- "Staccato"
}
}

# Forte
for (phrase in 1:length(ls_forte_2)){
for (note in 1:length(ls_forte_2[[phrase]])){
  dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_forte_2[[phrase]][note]] <- "Forte"
}
}

# Piano
for (phrase in 1:length(ls_piano_2)){
for (note in 1:length(ls_piano_2[[phrase]])){
  dt_vel$Subcomponent[dt_vel$Skill == "dynamics" & dt_vel$Note == ls_piano_2[[phrase]][note]] <- "Piano"
}
}

vel <- dt_vel[Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(Instance, Condition, Skill, Subcomponent)]
vel_all <- vel[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Subcomponent)]
```

## Each instance
```{r vel_stats_ind, echo = FALSE}
vel
```

## Averaged
```{r vel_stats_all, echo = FALSE}
vel_all
```

## Plots
```{r vel_plot, echo = FALSE}
p_vel <- ggplot(vel, aes(x = Instance, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
facet_grid(Subcomponent ~ .) +
labs(y = "Key Velocity", subtitle = "KV: Each instance (Dynamics)")
p_vel

p_vel_all <- ggplot(vel_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
labs(y = "Key Velocity", subtitle = "KV: Averaged (Dynamics)")
p_vel_all
```

# KV-Diff
```{r vel_diff, include = FALSE}
# calculate differences and remove the first note
dt_vel$Diff <- diff(c(0, dt_vel$Velocity))
dt_vel_diff <- dt_vel[Note != 1]
dt_vel$Diff <- NULL # remove Diff from dt_vel

# assign Interval
dt_vel_diff$Interval <- rep(1:66, nrow(dt_vel_diff)/66)

# assign Subcomponents
# for intervals
dt_vel_diff$Subcomponent <- NA
# Legato
for (phrase in 1:length(ls_legato)){
for (note in 1:length(ls_legato[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "articulation" & dt_vel_diff$Interval == ls_legato[[phrase]][note]] <- "Legato"
}
}
# Staccato
for (phrase in 1:length(ls_staccato)){
for (note in 1:length(ls_staccato[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "articulation" & dt_vel_diff$Interval == ls_staccato[[phrase]][note]] <- "Staccato"
}
}

# Forte
for (phrase in 1:length(ls_forte)){
for (note in 1:length(ls_forte[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == ls_forte[[phrase]][note]] <- "Forte"
}
}
# Piano
for (phrase in 1:length(ls_piano)){
for (note in 1:length(ls_piano[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == ls_piano[[phrase]][note]] <- "Piano"
}
}

# Assign Skill Change
for (i in change_1){
dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "articulation" & dt_vel_diff$Interval == i] <- "LtoS"
dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == i] <- "FtoP"
}
for (i in change_2){
dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "articulation" & dt_vel_diff$Interval == i] <- "StoL"
dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == i] <- "PtoF"
}

vel_diff <- dt_vel_diff[Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(Instance, Condition, Skill, Subcomponent)]
vel_diff_all <- vel_diff[, .(N = .N, Mean = mean(Mean), SD = sd(Mean)), by = .(Condition, Skill, Subcomponent)]
```

## Each instance
```{r vel__diff_stats_ind, echo = FALSE}
vel_diff
```

## Averaged
```{r vel_diff_stats_all, echo = FALSE}
vel_diff_all
```

## Plots
```{r vel_diff_plot, echo = FALSE}
p_vel_diff <- ggplot(vel_diff, aes(x = Instance, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
facet_grid(Subcomponent ~ .) +
labs(y = "KV difference", subtitle = "KV-Diff: Each instance (Dynamics)")
p_vel_diff

p_vel_diff_all <- ggplot(vel_diff_all, aes(x = Subcomponent, y = Mean, fill = Condition)) +
geom_bar(stat = "identity", position = position_dodge()) +
geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD),
              width = .2, position = position_dodge(.9)) +
labs(y = "KV difference", subtitle = "KV-Diff: Averaged (Dynamics)")
p_vel_diff_all
```