### 2021-01-25 analysis/stimuli/stimuli_analysis.Rmd
stimuli analysis from actual midfiles
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception: Stimuli Analysis)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

I chose stimuli with the IOIs within 2sd randomly from each condition (teaching, performing), each technique (articulation, dynamics).

Note: SubNr 8 was excluded due to an experimental error, SubNr 16 was excluded due to a deviated tempo (see analysis for [music-teaching: experiment-1](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-1))

# DVs
- IOIs (average tempo)
- Transition points
- CV (tempo variability)
- KOT (articulation)
- KV (dynamics)
- KV-diff (dynamics)

```{r setup, include = FALSE}
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
if (!require("ggplot2")) {install.packages("ggplot2"); require("ggplot2")}
```

```{r file, include = FALSE}
# create a list of data file names
lf <- list.files("../../experiment/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../../experiment/mid/", file, sep = ""))
   current$midfile <- file
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:134, nrow(data)/134)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity")
data$Condition <- "NA"
data$Skill <- "NA"
data[midfile %like% "_t_"]$Condition <- "teaching"
data[midfile %like% "_p_"]$Condition <- "performing"
data[startsWith(midfile, "a_")]$Skill <- "articulation"
data[startsWith(midfile, "d_")]$Skill <- "dynamics"

# numbering
instance = 1
data$Instance <- 0
for (file in unique(data$midfile)){
   data[midfile == file]$Instance <- instance
   instance = instance + 1
}

# define subcomponents
# For intervals
ls_legato <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_staccato <- list(c(9:15), c(25:31), c(34:40), c(50:56))
ls_forte <- list(c(1:7), c(17:23), c(42:48), c(58:64))
ls_piano <- list(c(9:15), c(25:31), c(34:40), c(50:56))

# For each note (velocity only)
ls_legato_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_staccato_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))
ls_forte_2 <- list(c(1:8), c(17:24), c(42:49), c(58:65))
ls_piano_2 <- list(c(9:16), c(25:32), c(34:41), c(50:57))

# Define Skill Change (LtoS, FtoP)
change_1 <- c(8, 24, 49)
# Define Skill Change (StoL, PtoF)
change_2 <- c(16, 41, 57)
```

# IOIs
```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:66, nrow(dt_ioi)/66)

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}
```

# All
## Articulation
```{r ioi-all-art, echo = FALSE}
ioi_art <- dt_ioi[Skill == "articulation"]

```

## Dynamics
```{r ioi-all-dyn, echo = FALSE}
ioi_dyn <- dt_ioi[Skill == "dynamics"]

```