### 2021-05-10 analysis/analysis.Rmd
upload analysis
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception 2: Analysis"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

- 15 participants were included for analysis.

# Experiment

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-2](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-2)).
- 36 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 72 stimuli). The half of them was selected from the teaching condition and the other half was from the performing condition.
  + 18 for teaching-articulation performances
  + 18 for performing-articulation performances
  + 18 for teaching-dynamics performances
  + 18 for performing-dynamics performances

## Design

- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques (expressions) separately. The order of the blocks was counterbalanced across participants.
- 36 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented across participants.

## Procedure
1. Musicians (> 6yo experience) were asked to listen to a number of recordings and judge whether each performance was produced for teaching purposes or not.

Actual instruction:
```
Each performance was produced in order to either 1) teach the musical expressive technique (as a teacher) or 2) perform their best (as a performer).

You will be asked to judge whether each performer had the intention to teach or not by pressing the 'Yes' <Left> or 'No' <Right> key.
```

2. Participants pressed "YES" if they think the recording was produced for teaching (as a teacher) whereas they pressed "NO" if they thought the recording was produced for performing (as a performer).

3. Correct answers would be 1) participants pressed "YES" when they listened to teaching performances or 2) participants pressed "NO" when they listened to performing performances.

## Variables
- IOIs (tempo)
- IOIs at transition points (tempo)
- CV (tempo)
- KOT (articulation)
- KV (dynamics)
- KV-Diff (dynamics)

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("qqplotr")) {install.packages("qqplotr"); require("qqplotr")}
# stats
if (!require("ppcor")) {install.packages("ppcor"); require("ppcor")}
if (!require("performance")) {install.packages("performance"); require("performance")}
```

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("data")
combined <- lapply(data_ls, function(f){
  fread(paste("data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)

# remove irrelevant strings
data <- all_data[expMode == "experiment"]
data$MidFile <- gsub("./mid/", "", data$midFile)
data$MidFile <- gsub(".mid", "", data$MidFile)

# name conditions
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", midFile)]$Condition <- "teaching"
data[grepl("_p_", midFile)]$Condition <- "performing"
data[grepl("a_", midFile)]$Skill <- "articulation"
data[grepl("d_", midFile)]$Skill <- "dynamics"
```

# Judged as "teaching" (%) for each stimulus
```{r teaching, echo = FALSE}
data$Teaching <- 0
data[rating == "Yes"]$Teaching <- 1

# individual
individual <- data[subjectNumber != 4, .(N = .N, Sum = sum(Teaching)), by = .(subjectNumber, Condition, Skill, MidFile)]
individual<- individual[order(subjectNumber)]

# all
all <- individual[, .(N = .N, Mean = mean(Sum), SD = sd(Sum), Mean_Percent = (mean(Sum))*100, SD_Percent = (sd(Sum))*100), by = .(Condition, Skill, MidFile)]
all <- all[order(MidFile)]
all
```

```{r midi, include = FALSE}
# create a list of data file names
lf <- list.files("../experiment/mid/", pattern = "mid")

# combine data
all_data <- data.table()
for (file in lf){
   current <- readMidi(paste("../experiment/mid/", file, sep = ""))
   current$MidFile <- gsub(".mid", "", file)
   all_data <- rbind(all_data, current)
}

data <- all_data[event == "Note On" | event == "Note Off"]
data$NoteNr <- rep(1:144, nrow(data)/144)
data$channel <- NULL
data$parameterMetaSystem <- NULL
data$track <- NULL
data$type <- NULL
data$time2 <- (data$time/960)*1000 # deltatime to time (ms)

# labelling
colnames(data)[c(3,4)] <- c("Pitch", "Velocity") # time event also change
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", MidFile)]$Condition <- "teaching"
data[grepl("_p_", MidFile)]$Condition <- "performing"
data[grepl("a_", MidFile)]$Skill <- "articulation"
data[grepl("d_", MidFile)]$Skill <- "dynamics"

# define subcomponents
# for intervals
ls_legato <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_staccato <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))
ls_forte <- list(c(1:4), c(9:16), c(21:24), c(40:46))
ls_piano <- list(c(6:7), c(18:19), c(29:31), c(36:38), c(48:50), c(53:55), c(58:64))

# for each note (velocity)
ls_legato_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_staccato_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))
ls_forte_2 <- list(c(1:5), c(9:17), c(21:26), c(40:47))
ls_piano_2 <- list(c(6:8), c(18:20), c(29:32), c(36:39), c(48:51), c(53:56), c(58:65))

# define Component Change (LtoS, FtoP)
change_1 <- c(5, 17, 47)
# define Component Change (StoL, PtoF)
change_2 <- c(8, 20, 39)
```

# Results
- significance levels: `***: 0.001, **: 0.01, *: 0.05`

# 1. Correlations
- Summaries of pure correlations between each parameter and the teaching judgement (**without controlling for potential third variables**)

# IOIs
- average tempo and teaching judgement
- linear regression lines and 95% confidence interval (gray)

```{r ioi, include = FALSE}
data_onset <- data[event == "Note On"]
data_offset <- data[event == "Note Off"]

data_onset$IOI <- diff(c(0, data_onset$time2))
dt_ioi <- data_onset[NoteNr != 1]
dt_ioi$Interval <- rep(1:71, nrow(dt_ioi)/71)

# assign Subcomponents
dt_ioi$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
 for (note in 1:length(ls_legato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
 }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
 for (note in 1:length(ls_staccato[[phrase]])){
   dt_ioi[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
 }
}

# Forte
for (phrase in 1:length(ls_forte)){
 for (note in 1:length(ls_forte[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
 }
}
# Piano
for (phrase in 1:length(ls_piano)){
 for (note in 1:length(ls_piano[[phrase]])){
   dt_ioi[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
 }
}

# assign Subcomponent Change
for (number in change_1){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
 dt_ioi[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
 dt_ioi[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}
```

## All (n.s.)
```{r ioi-all, echo = FALSE}
rating <- all # rating data

ioi <- dt_ioi[Subcomponent != "NA", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(MidFile, Skill)]

# sorted
ioi <- ioi[order(MidFile)]

# plot
ioi$Teaching <- rating$Mean_Percent
ioi$TeachingSD <- rating$SD_Percent

ggscatter(ioi, x = "Mean", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Articulation (n.s.)
```{r, echo = FALSE}
cor.test(ioi[Skill == "articulation"]$Teaching, ioi[Skill == "articulation"]$Mean)
```

## Dynamics (n.s.)
```{r, echo = FALSE}
cor.test(ioi[Skill == "dynamics"]$Teaching, ioi[Skill == "dynamics"]$Mean)
```

# Tempo at transition points
- tempo at transition points (e.g., legato to staccato / forte to piano) and teaching judgement
- linear regression line (black) and 95% confidence interval (gray)

## All (n.s.)
```{r ioi-tra-all, echo = FALSE}
ioi_tra <- dt_ioi[Subcomponent == "LtoS" | Subcomponent == "StoL" | Subcomponent == "FtoP" | Subcomponent == "PtoF", .(N = .N, Mean = mean(IOI), SD = sd(IOI)), by = .(MidFile, Skill)]

# sorted
ioi_tra <- ioi_tra[order(MidFile)]

# plot
ioi_tra$Teaching <- rating$Mean_Percent
ioi_tra$TeachingSD <- rating$SD_Percent

ggscatter(ioi_tra, x = "Mean", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "IOIs (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Articulation (n.s.)
```{r, echo = FALSE}
cor.test(ioi_tra[Skill == "articulation"]$Teaching, ioi_tra[Skill == "articulation"]$Mean)
```

## Dynamics (n.s.)
```{r, echo = FALSE}
cor.test(ioi_tra[Skill == "dynamics"]$Teaching, ioi_tra[Skill == "dynamics"]$Mean)
```

# CV (tempo variability)
- tempo variability and teaching judgement
- linear regression lines and 95% confidence interval (gray)

## All (n.s.)
```{r cv, echo = FALSE}
cv <- dt_ioi[Subcomponent != "NA", .(N = .N, CV = sd(IOI)/mean(IOI)), by = .(MidFile, Skill)]

# sorted
cv <- cv[order(MidFile)]

# plot
cv$Teaching <- rating$Mean_Percent
cv$TeachingSD <- rating$SD_Percent

ggscatter(cv, x = "CV", y = "Teaching", color = "Skill", add = "reg.line",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = TRUE,
          xlab = "CV", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Articulation (n.s.)
```{r, echo = FALSE}
cor.test(cv[Skill == "articulation"]$Teaching, cv[Skill == "articulation"]$CV)
```

## Dynamics (n.s.)
```{r, echo = FALSE}
cor.test(cv[Skill == "dynamics"]$Teaching, cv[Skill == "dynamics"]$CV)
```

# KOT
- articulation and teaching judgement
- linear regression lines and 95% confidence interval (gray)

```{r kot, include = FALSE}
data_onset$KOT <- 0
for (row in 1:nrow(data_onset)){
   if (row < nrow(data_onset)){
      data_onset$KOT[row+1] <- data_offset$time2[row] - data_onset$time2[row+1] # offset(n) - onset(n+1)
   }
}

dt_kot <- data_onset[NoteNr != 1]
dt_kot$Interval <- rep(1:71, nrow(dt_kot)/71)

# assign Subcomponents
dt_kot$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
   for (note in 1:length(ls_legato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato)){
   for (note in 1:length(ls_staccato[[phrase]])){
     dt_kot[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte)){
   for (note in 1:length(ls_forte[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_forte[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano)){
   for (note in 1:length(ls_piano[[phrase]])){
     dt_kot[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
   }
}
# assign Subcomponent Change
for (number in change_1){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "LtoS"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "FtoP"
}
for (number in change_2){
   dt_kot[Skill == "articulation" & Interval == number]$Subcomponent <- "StoL"
   dt_kot[Skill == "dynamics" & Interval == number]$Subcomponent <- "PtoF"
}
```

## All
```{r kot-all, echo = FALSE}
kot_all <- dt_kot[Subcomponent == "Legato" | Subcomponent == "Staccato" | Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(KOT), SD = sd(KOT)), by = .(MidFile, Skill, Subcomponent)]

# sorted
kot_all <- kot_all[order(MidFile)]

# plot
kot_all$Teaching <- rep(rating$Mean_Percent, each = 2)
kot_all$TeachingSD <- rep(rating$SD_Percent, each = 2)

ggscatter(kot_all, x = "Mean", y = "Teaching", color = "Subcomponent", add = "reg.line", facet.by = "Skill",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = FALSE,
          xlab = "KOT (ms)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```
## Legato (n.s.)
```{r, echo = FALSE}
cor.test(kot_all[Subcomponent == "Legato"]$Teaching, kot_all[Subcomponent == "Legato"]$Mean)
```

## Staccato (n.s.)
```{r, echo = FALSE}
cor.test(kot_all[Subcomponent == "Staccato"]$Teaching, kot_all[Subcomponent == "Staccato"]$Mean)
```

## Forte (n.s.)
```{r, echo = FALSE}
cor.test(kot_all[Subcomponent == "Forte"]$Teaching, kot_all[Subcomponent == "Forte"]$Mean)
```

## Piano (n.s.)
```{r, echo = FALSE}
cor.test(kot_all[Subcomponent == "Piano"]$Teaching, kot_all[Subcomponent == "Piano"]$Mean)
```

# KV
- dynamics and teaching judgement
- linear regression lines and 95% confidence interval (gray)

```{r vel, include = FALSE}
dt_vel <- data_onset
dt_vel$Note <- rep(1:72, nrow(dt_vel)/72)
# assign Subcomponents
# for each note
dt_vel$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato_2)){
   for (note in 1:length(ls_legato_2[[phrase]])){
     dt_vel[Skill == "articulation" & Note == ls_legato_2[[phrase]][note]]$Subcomponent <- "Legato"
   }
}
# Staccato
for (phrase in 1:length(ls_staccato_2)){
   for (note in 1:length(ls_staccato_2[[phrase]])){
     dt_vel[Skill == "articulation" & Note == ls_staccato_2[[phrase]][note]]$Subcomponent <- "Staccato"
   }
}
# Forte
for (phrase in 1:length(ls_forte_2)){
   for (note in 1:length(ls_forte_2[[phrase]])){
     dt_vel[Skill == "dynamics" & Note == ls_forte_2[[phrase]][note]]$Subcomponent <- "Forte"
   }
}
# Piano
for (phrase in 1:length(ls_piano_2)){
   for (note in 1:length(ls_piano_2[[phrase]])){
     dt_vel[Skill == "dynamics" & Note == ls_piano_2[[phrase]][note]]$Subcomponent <- "Piano"
   }
}
```

## All
```{r vel-all, echo = FALSE}
vel_all <- dt_vel[Subcomponent == "Legato" | Subcomponent == "Staccato" | Subcomponent == "Forte" | Subcomponent == "Piano", .(N = .N, Mean = mean(Velocity), SD = sd(Velocity)), by = .(MidFile, Skill, Subcomponent)]

# sorted
vel_all <- vel_all[order(MidFile)]

# plot
vel_all$Teaching <- rep(rating$Mean_Percent, each = 2)
vel_all$TeachingSD <- rep(rating$SD_Percent, each = 2)

ggscatter(vel_all, x = "Mean", y = "Teaching", color = "Subcomponent", add = "reg.line", facet.by = "Skill",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = FALSE,
          xlab = "Velocity (0-127)", ylab = "Judged as teaching (%)") + ylim(0, 100)
```

## Forte (*)
```{r, echo = FALSE}
cor.test(vel_all[Subcomponent == "Forte"]$Teaching, vel_all[Subcomponent == "Forte"]$Mean)
```

## Piano (*)
```{r, echo = FALSE}
cor.test(vel_all[Subcomponent == "Piano"]$Teaching, vel_all[Subcomponent == "Piano"]$Mean)
```
## Legato (n.s.)
```{r, echo = FALSE}
cor.test(vel_all[Subcomponent == "Legato"]$Teaching, vel_all[Subcomponent == "Legato"]$Mean)
```

## Staccato (n.s.)
```{r, echo = FALSE}
cor.test(vel_all[Subcomponent == "Staccato"]$Teaching, vel_all[Subcomponent == "Staccato"]$Mean)
```

# KV-Diff
- dynamics changes at transition points (between forte and piano) and teaching judgement
- linear regression lines and 95% confidence interval (gray)

```{r vel-diff, include = FALSE}
data_onset$Diff <- diff(c(0, data_onset$Velocity))
dt_vel_diff <- data_onset[NoteNr != 1]
# assign Interval
dt_vel_diff$Interval <- rep(1:71, nrow(dt_vel_diff)/71)

# for intervals
dt_vel_diff$Subcomponent <- "NA"
# Legato
for (phrase in 1:length(ls_legato)){
for (note in 1:length(ls_legato[[phrase]])){
  dt_vel_diff[Skill == "articulation" & Interval == ls_legato[[phrase]][note]]$Subcomponent <- "Legato"
}
}
# Staccato
for (phrase in 1:length(ls_staccato)){
for (note in 1:length(ls_staccato[[phrase]])){
  dt_vel_diff[Skill == "articulation" & Interval == ls_staccato[[phrase]][note]]$Subcomponent <- "Staccato"
}
}
# Forte
for (phrase in 1:length(ls_forte)){
for (note in 1:length(ls_forte[[phrase]])){
  dt_vel_diff$Subcomponent[dt_vel_diff$Skill == "dynamics" & dt_vel_diff$Interval == ls_forte[[phrase]][note]] <- "Forte"
}
}
# Piano
for (phrase in 1:length(ls_piano)){
for (note in 1:length(ls_piano[[phrase]])){
  dt_vel_diff[Skill == "dynamics" & Interval == ls_piano[[phrase]][note]]$Subcomponent <- "Piano"
}
}
# Assign Skill Change
for (i in change_1){
dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "LtoS"
dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "FtoP"
}
for (i in change_2){
dt_vel_diff[Skill == "articulation" & Interval == i]$Subcomponent <- "StoL"
dt_vel_diff[Skill == "dynamics" & Interval == i]$Subcomponent <- "PtoF"
}
```

## All
```{r vel-diff-all, echo = FALSE}
vel_diff_all <- dt_vel_diff[Subcomponent == "FtoP" | Subcomponent == "PtoF" | Subcomponent == "LtoS" | Subcomponent == "StoL", .(N = .N, Mean = mean(Diff), SD = sd(Diff)), by = .(MidFile, Skill, Subcomponent)]

# sorted
vel_diff_all <- vel_diff_all[order(MidFile)]

# plot
vel_diff_all$Teaching <- rep(rating$Mean_Percent, each = 2)
vel_diff_all$TeachingSD <- rep(rating$SD_Percent, each = 2)

ggscatter(vel_diff_all, x = "Mean", y = "Teaching", color = "Subcomponent", add = "reg.line", facet.by = "Skill",
          add.params = list(fill = "lightgray"), conf.int = TRUE, cor.coef = FALSE,
          xlab = "Velocity Difference (-127-127)", ylab = "Judged as teaching (%)") +ylim(0, 100)
```

## Forte to Piano (***)
```{r, echo = FALSE}
cor.test(vel_diff_all[Subcomponent == "FtoP"]$Teaching, vel_diff_all[Subcomponent == "FtoP"]$Mean)
```

## Piano to Forte (***)
```{r, echo = FALSE}
cor.test(vel_diff_all[Subcomponent == "PtoF"]$Teaching, vel_diff_all[Subcomponent == "PtoF"]$Mean)
```
## Legato to Staccato (*)
```{r, echo = FALSE}
cor.test(vel_diff_all[Subcomponent == "LtoS"]$Teaching, vel_diff_all[Subcomponent == "LtoS"]$Mean)
```

## Staccato to Legato (n.s.)
```{r, echo = FALSE}
cor.test(vel_diff_all[Subcomponent == "StoL"]$Teaching, vel_diff_all[Subcomponent == "StoL"]$Mean)
```

# 2. Partical correlation (only for KOT, KV, KV-Diff)
- Control for third variables
- When investigating the effect of KOT, we control for IOIs and dynamics (KV, KV-Diff)
- When investigating the effect of KV or KV-Diff, we control for IOIs and articulation (KOT)

```{r partial, include = FALSE}
partial <- dt_ioi[Subcomponent != "NA", .(N = .N, IOI = mean(IOI), IOISD = sd(IOI)), by = .(MidFile, Skill)]
partial <- rbind(partial, partial)
partial <- partial[order(MidFile)]

partial$Subcomponent <- kot_all$Subcomponent
partial$KOT <- kot_all$Mean
partial$KOTSD <- kot_all$SD
partial$KV <- vel_all$Mean
partial$KVSD <- vel_all$SD
partial$Subcomponent2 <- vel_diff_all$Subcomponent
partial$KVDiff <- vel_diff_all$Mean
partial$KVDiffSD <- vel_diff_all$SD

partial$Teaching <- rep(rating$Mean_Percent, each = 2)
partial$TeachingSD <- rep(rating$SD_Percent, each = 2)
```

# KOT
## Legato (n.s.)
```{r kot-partial-leg, echo = TRUE}
pcor.test(partial[Subcomponent == "Legato"]$KOT, partial[Subcomponent == "Legato"]$Teaching, partial[Subcomponent == "Legato", c("IOI", "KV", "KVDiff")])
```

## Staccato (n.s.)
```{r kot-partial-sta, echo = TRUE}
pcor.test(partial[Subcomponent == "Staccato"]$KOT, partial[Subcomponent == "Staccato"]$Teaching, partial[Subcomponent == "Staccato", c("IOI", "KV", "KVDiff")])
```

# KV
## Forte (n.s.)
```{r vel-partial-for, echo = TRUE}
pcor.test(partial[Subcomponent == "Forte"]$KV, partial[Subcomponent == "Forte"]$Teaching, partial[Subcomponent == "Forte", c("IOI", "KOT", "KVDiff")])
```

## Piano (n.s.)
```{r vel-partial-pia, echo = TRUE}
pcor.test(partial[Subcomponent == "Piano"]$KV, partial[Subcomponent == "Piano"]$Teaching, partial[Subcomponent == "Piano", c("IOI", "KOT", "KVDiff")])
```

# KV-Diff
## Forte to Piano (***)
```{r vel-diff-partial-ftop, echo = TRUE}
pcor.test(partial[Subcomponent2 == "FtoP"]$KVDiff, partial[Subcomponent2 == "FtoP"]$Teaching, partial[Subcomponent2 == "FtoP", c("IOI", "KOT", "KV")])
```

## Piano to Forte (***)
```{r vel-diff-partial-ptof, echo = TRUE}
pcor.test(partial[Subcomponent2 == "PtoF"]$KVDiff, partial[Subcomponent2 == "PtoF"]$Teaching, partial[Subcomponent2 == "PtoF", c("IOI", "KOT", "KV")])
```

# Multiple regression
Note: additive - **no interaction considered**

## Legato
```{r, echo = TRUE}
m1 <- lm(Teaching ~ IOI + KOT + KV + KVDiff, data = partial[Subcomponent == "Legato"])
summary(m1)
check_model(m1)
```

## Staccato
```{r, echo = TRUE}
m2 <- lm(Teaching ~ IOI + KOT + KV + KVDiff, data = partial[Subcomponent == "Staccato"])
summary(m2)
check_model(m2)
```

## Forte
```{r, echo = TRUE}
m3 <- lm(Teaching ~ IOI + KOT + KV + KVDiff, data = partial[Subcomponent == "Forte"])
summary(m3)
check_model(m3)
```

## Piano
```{r, echo = TRUE}
m4 <- lm(Teaching ~ IOI + KOT + KV + KVDiff, data = partial[Subcomponent == "Piano"])
summary(m4)
check_model(m4)
```