### 2022-03-14 experiment-1/analysis/prereg_analysis/prereg_analysis.Rmd
update analysis for first draft
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

---
title: "Music Perception 1: Preregistered Analysis"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

- 21 participants (experimental error = 1) >> 20 participants were included for analysis.

# Experiment
## Preregistered Analysis
The preregistered analysis aims to investigate whether participants could distinguish performances produced for teaching purposes from those produced for performing purposes. For more detailed explanations, please see [the OSF preregistration](https://osf.io/z3j69).

## Stimuli
- Stimuli were selected from the previous study ([the sound of teaching music/experiment-1](https://github.com/atsukotominaga/music-teaching/tree/main/experiment-1)).
- 48 stimuli were randomly chosen from each technique (either articulation or dynamics - therefore in total 96 stimuli). Half of them was selected from the teaching condition and the other half was from the performing condition.
  + 24 for teaching-articulation performances
  + 24 for performing-articulation performances
  + 24 for teaching-dynamics performances
  + 24 for performing-dynamics performances

## Design
- 2 blocks (articulation / dynamics): participants were asked to do the same task for two techniques separately. The order of the blocks was counterbalanced across participants.
- 48 trials for each block (each stimulus was presented only once) and the stimuli were randomly presented within a block.

## Procedure
1. Musicians (> 6yo experience) were asked to listen to a number of recordings and judge whether each performance was produced for teaching purposes or not.

Actual instruction:
```
Each performance was produced in order to either 1) teach the musical expressive technique (as a teacher) or 2) perform their best (as a performer).

You will be asked to judge whether each performer had the intention to teach or not by pressing the 'Yes' <Left> or 'No' <Right> key.
```

2. Participants pressed "YES" if they think the recording was produced for teaching (as a teacher) whereas they pressed "NO" if they thought the recording was produced for performing (as a performer).

3. Correct answers would be 1) participants pressed "YES" when they listened to teaching performances or 2) participants pressed "NO" when they listened to performing performances.

## Dependent variables
- Accuracy (Correct %)
   + "YES" to teaching stimuli: correct
   + "NO" to performing stimuli: correct
   + "NO" to teaching stimuli: incorrect
   + "YES" to performing stimuli: incorrect

```{r packages, include = FALSE}
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# midi
if (!require("tuneR")) {install.packages("tuneR"); require("tuneR")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
# stats
if (!require("afex")) {install.packages("afex"); require("afex")}
if (!require("rstatix")) {install.packages("rstatix"); require("rstatix")}
```

```{r read, include = FALSE}
# read files and combine them
data_ls <- list.files("../../data", pattern = "csv")
combined <- lapply(data_ls, function(f){
  fread(paste("../../data/", f, sep = ""), sep = ",")
})
all_data <- do.call(rbind.data.frame, combined)

# remove irrelevant strings
data <- all_data[expMode == "experiment"]
data$MidFile <- gsub("./mid/", "", data$midFile)
data$MidFile <- gsub(".mid", "", data$MidFile)

# name conditions
data$Condition <- "NA"
data$Skill <- "NA"
data[grepl("_t_", midFile)]$Condition <- "teaching"
data[grepl("_p_", midFile)]$Condition <- "performing"
data[grepl("a_", midFile)]$Skill <- "articulation"
data[grepl("d_", midFile)]$Skill <- "dynamics"

# check correct or incorrect
data$Correct <- 0
data[Condition == "teaching" & rating == "Yes"]$Correct <- 1
data[Condition == "performing" & rating == "No"]$Correct <- 1

all <- data[expMode == "experiment", .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
all[order(subjectNumber, Condition, Skill)]
```

# Results
- significance levels: `***: 0.001, **: 0.01, *: 0.05`

## 1. Statistics
## All
```{r acc, echo = FALSE}
# individual
ind_data <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber)]
ind_data <- ind_data[order(subjectNumber)]

# all
subset <- ind_data[, .(N = .N, Mean_Percent = mean(Percent), SD_Percent = sd(Percent))]
subset
```

```{r plot, echo = FALSE}
ggboxplot(ind_data, x = "N", y = "Percent", add = "dotplot", xlab = FALSE) + ylab("Correct %") + theme(axis.text.x = element_blank()) + geom_hline(yintercept = 50, linetype = "dashed")
```

### One-sample *t*-test
```{r t-test, echo = FALSE}
# normality check
shapiro.test(ind_data$Percent)

ttest_all <- t.test(ind_data$Percent, mu = 50, alternative = "two.sided")
ttest_all
```

### Effect size
```{r, echo = FALSE}
coheneff_all <- ind_data %>% cohens_d(Percent ~ 1, mu = 50)
coheneff_all
```

## Articulation vs. Dynamics
Reference for a paired plot with jitter: [https://stackoverflow.com/questions/48042810/add-jitter-to-ggpaired#48044022](https://stackoverflow.com/questions/48042810/add-jitter-to-ggpaired#48044022)

```{r plot2, echo = FALSE}
ind_data_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Skill)]
ind_data_ad <- ind_data_ad[order(subjectNumber, Skill)]

subset_ad <- ind_data_ad[, .(N = .N, Mean_Percent = mean(Percent), SD_Percent = sd(Percent)), by = .(Skill)]
subset_ad

data_box <- data.table(subjectNumber = ind_data_ad[Skill == "articulation"]$subjectNumber, articulation = ind_data_ad[Skill == "articulation"]$Percent, dynamics = ind_data_ad[Skill == "dynamics"]$Percent)

# plot with jitter
n <- length(data_box$articulation)
d <- data.frame(y = c(data_box$articulation, data_box$dynamics), 
                x = rep(c(1, 2), each = n),
                id = factor(rep(1:n, 2)))

set.seed(321)    
d$xj <- jitter(d$x, amount = .05)
ggplot(data = d, aes(y=y)) +
  geom_boxplot(aes(x = x, group = x), width = 0.5, outlier.shape = NA) +
  geom_point(aes(x = xj)) +
  geom_line(aes(x = xj, group = id), color = "gray", size = .4) + labs(x = "Skill", y = "Correct %") +
  scale_x_continuous(breaks = c(1, 2), labels = c("articulation", "dynamics"), limits = c(0.5, 2.5)) + geom_hline(yintercept = 50, linetype = "dashed") + theme_pubr()

# plot without jitter
# ggpaired(data_box, cond1 = "articulation", cond2 = "dynamics", line.color = "gray", line.size = 0.4, xlab = "Skill", ylab = "Correct %") + geom_hline(yintercept = 50, linetype = "dashed")
```

### One-sample *t*-test
### Articulation
```{r t-test-a, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "articulation"]$Percent)

ttest_art <- t.test(ind_data_ad[Skill == "articulation"]$Percent, mu = 50, alternative = "two.sided")
ttest_art
```

### Effect size
```{r, echo = FALSE}
coheneff_art <- ind_data_ad[Skill == "articulation"] %>% cohens_d(Percent ~ 1, mu = 50)
coheneff_art
```

### Dynamics
```{r t-test-d, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent)

ttest_dyn <- t.test(ind_data_ad[Skill == "dynamics"]$Percent, mu = 50, alternative = "two.sided")
ttest_dyn
```

### Effect size
```{r, echo = FALSE}
coheneff_dyn <- ind_data_ad[Skill == "dynamics"] %>% cohens_d(Percent ~ 1, mu = 50)
coheneff_dyn
```

### Paired *t*-test between articulation and dynamics
```{r p-t-test, echo = FALSE}
# normality check
shapiro.test(ind_data_ad[Skill == "dynamics"]$Percent-ind_data_ad[Skill == "articulation"]$Percent)

ttest_paired <- t.test(ind_data_ad[Skill == "articulation"]$Percent, ind_data_ad[Skill == "dynamics"]$Percent, paired = TRUE)
ttest_paired
```
### Effect size
```{r, echo = FALSE}
coheneff_paired <- ind_data_ad %>% cohens_d(Percent ~ Skill, mu = 50)
coheneff_paired
```

## Two-way ANOVA (Condition * Skill)
```{r plot3, echo = FALSE}
ind_data_tp_ad <- data[subjectNumber != 9, .(N = .N, Sum = sum(Correct), Percent = (sum(Correct)/.N)*100), by = .(subjectNumber, Condition, Skill)]
ind_data_tp_ad <- ind_data_tp_ad[order(subjectNumber, Condition, Skill)]

ggboxplot(ind_data_tp_ad, "Skill", "Percent", color = "Condition") + geom_hline(yintercept = 50, linetype = "dashed")
```

```{r, echo = FALSE}
anova <- aov_ez("subjectNumber", "Percent",
                data = ind_data_tp_ad, within = c("Condition", "Skill"))
anova
```

```{r export, include = FALSE}
knitr::purl("prereg_analysis.Rmd")
```