### 2022-10-14 experiment-2/analysis/questionnaire.Rmd

---
title: "Music Perception: Questionnaire (Experiment 2)"
output: html_notebook
---

- Last checked: `r format(Sys.Date(), "%d-%b-%Y")`

```{r setup, include = FALSE}
# packages
# data manipulation
if (!require("data.table")) {install.packages("data.table"); require("data.table")}
# plot
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
```

# Questionnaire

```{r read, echo = FALSE}
dt <- fread("questionnaire.csv", header = T, sep = ",", dec = ".", na.string = "NA")

# change some colnames
colnames(dt)[2] <- "SubNr"
colnames(dt)[3] <- "Age"
colnames(dt)[4] <- "Gender"
colnames(dt)[7] <- "Handedness"

# extract the total number of years of training for the longest instrument (input manually based on the original csv file)
dt$Training <- c(14, 13, 7, 18, 23, 12, 17, 20, 12, 7, 12, 7, 7, 13, 20, 15, 6, 15, 8, 5, 20)

# exclude participants
dt_included <- dt[SubNr != 4]

print(dt_included)
```

## Descriptive stats for included participants
- Number of included participants: `r nrow(dt_included)` (mean age: `r round(mean(dt_included$Age), 2)`, sd: `r round(sd(dt_included$Age), 2)`)
- Male: `r nrow(dt_included[Gender == "Male"])` (mean age: `r round(mean(dt_included[Gender == "Male"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Male"]$Age), 2)`)
- Female: `r nrow(dt_included[Gender == "Female"])` (mean age: `r round(mean(dt_included[Gender == "Female"]$Age), 2)`, sd: `r round(sd(dt_included[Gender == "Female"]$Age), 2)`)
- Average training years: `r round(mean(dt_included$Training), 2)` years (*SD* = `r round(sd(dt_included$Training), 2)`)
- Handedness: Right (`r nrow(dt_included[Handedness == "Right"])`), Left (`r nrow(dt_included[Handedness == "Left"])`), Both (`r nrow(dt_included[Handedness == "Both"])`)

## Some descriptive questions
1. Did you try to perform differently when you are asked to play as a teacher during the experiment? If so, please describe how you changed your performance?

```{r 1, echo = FALSE}
data.table("Answer" = dt_included$`Have you noticed anything special regarding the tasks in the experiment? If any, please describe below.`)
```

```{r export, include = FALSE}
knitr::purl("questionnaire.Rmd")
```